---
title: "Hart Island Lobster Traps"
author: "Big Harts"
output: html_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(leaflet)
library(sf)
library(dplyr)
```

## 1. Introduction

This data set was provided from the Rozalia Project, a non-profit that works on collection, research, and education of marine debis around the Gulf of Maine. Most of what debris is collected and cataloged can be tied to the fishing industries in the region. Collecting lobster trap data through these methods has the potential to to help this industry understand trap losses in this region in order to minimize/prevent future loss of these mataerials.    

This data was collected on August 24th, 2023 on Hart Island, a small island just
Southwest of Port Clyde, ME. Volunteers worked to remove lost, abandoned and derelict fishing gear and other coastal debris. For each lobster trap found, the trap tags were pulled and recorded including: tag count, data collected (MDY), number of tags in bundle, state, zone, area, tag ID, trap batch number, tag color, shoreline where found, GPS (lat/long), tag year, additional info on tag, named individual, and home port. 

With this data we aim to explore the relationship between abandoned lobster traps and the individuals who deployed them and the lifespan of traps. We hope that this data can be added to a more substantial dataset including more islands across the Gulf of Maine as more are surveyed. 

## 2. Data

```{r read_lobsters}
lobsters <- read_csv(file = "../data/lobsters.csv")
glimpse(lobsters)
```

Remove unnecessary text rows at bottom of dataset:
```{r glimpse_true_lobsters}
true_lobsters <- head(lobsters, -4)
glimpse(true_lobsters)
```

## 3. Data analysis plan

1) We hope to use tag year and color, # of tags in bundle, named individual, tag ID number, and homeport.

2) We will try to find data showing the # of traps per individual so as to better understand the ratio of lost traps on Hart island to total # of traps. In addition, we hope to compare the home port of the traps to the number of tags in each bundle to see if there's a link between the longevity of a trap and the port it came from. 

3) The prelim_vis_testing graph below helps show the number of lost traps per individual, and the age_traps graph shows the correlation between age of traps and number of traps.

4) We hope to build on the graphs below to better show relationships between traps and individuals and show the range of trap age. 

5) We are also considering looking to see if there is a correlation between a trap's homeport and the number of tags in the bundle (which represents how many )

```{r prelim_vis_testing}
ggplot(data = true_lobsters, mapping = aes(x = `NAMED INDIVIDUAL`)) +
  geom_bar() +
theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


```{r trap-age-graph}
ggplot(data = true_lobsters, mapping = aes(x = `# TAGS IN BUNDLE`)) +
  geom_bar(fill = "darkblue") +
  labs(x = "Age of Traps (Number of Tags in Bundle)", 
       y = "Number of Traps",
       title = "Age of Traps Collected on Hart Island") +
  theme_linedraw() +
  scale_x_continuous(breaks = seq(0, 10, by = 1))
```
<<<<<<< HEAD
```{r tag-ID-&-port}
trap_data <- true_lobsters %>%
=======
```{r}
  trap_data <- true_lobsters %>%
>>>>>>> ad40b5bb0c2f8a4fd096b39e39fb5ee82ef8c706
  count(`TAG ID NUMBER`, `HOME PORT`) %>%
  arrange(desc(n)) %>%
  mutate(`HOME PORT` = if_else(`HOME PORT` == "N/A", "UNKNOWN", `HOME PORT`)) 
         

trap_data$`TAG ID NUMBER` <- fct_reorder(trap_data$`TAG ID NUMBER`, trap_data$n, .desc = TRUE)

ggplot(data = trap_data, aes(x = `TAG ID NUMBER`, y = n, fill = `HOME PORT`)) +
  geom_bar(stat = "identity") +
  labs(x = "Tag ID Number", y = "Number of Traps",
       title = "Trap IDs Found on Hart Island") +
  theme_linedraw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_viridis_d(option = "H", name = "Home Port", direction = -1)


```


```


```{r homeport_coords}
homeport_coords <- true_lobsters |>
  mutate(
    "HOMEPORT LAT" = case_when(
     `HOME PORT` == "PORT CLYDE" ~ 43.9271,
     `HOME PORT` == "WINTER HBR" ~ 44.3948,
     `HOME PORT` == "ROCKLAND" ~ 44.1037,
     `HOME PORT` == "TENANTS HBR" ~ 43.9682,
     `HOME PORT` == "CUSHING" ~ 44.0138,
     `HOME PORT` == "OWLS HEAD" ~ 44.0823,
     `HOME PORT` == "ST GEORGE" ~ 44.0165
    )) |>
  mutate(
    "HOMEPORT LONG" = case_when(
     `HOME PORT` == "PORT CLYDE" ~ -69.2583,
     `HOME PORT` == "WINTER HBR" ~ -68.0830,
     `HOME PORT` == "ROCKLAND" ~ -69.1089,
     `HOME PORT` == "TENANTS HBR" ~ -69.2088,
     `HOME PORT` == "CUSHING" ~ -69.2615,
     `HOME PORT` == "OWLS HEAD" ~ -69.0573,
     `HOME PORT` == "ST GEORGE" ~ -69.1989
    ))
```


```{r anon_data}
lobsters_anon <- homeport_coords |>   
  mutate(
    "INDIVIDUALS" = case_when(
      `NAMED INDIVIDUAL` == "THOMAS YOUNG/LC2O/WINTER HBR" ~ "T.Y.",
      `NAMED INDIVIDUAL` == "DOUGLAS ANDERSON/LC3/PORT CLYDE" ~ "D.A.",
      `NAMED INDIVIDUAL` == "KEITH YORK/LC2/ROCKLAND" ~ "K.Y.",
      `NAMED INDIVIDUAL` == "PETER HENDERSON/LNC/TENANTS HBR" ~ "P.H.",                      
      `NAMED INDIVIDUAL` == "CHRISTOPHER ANDERSON/LC2/PORT CLYDE" ~ "C.A.",                  
      `NAMED INDIVIDUAL` == "GARY LIBBY/LC2/PORT CLYDE" ~ "G.L.",                            
      `NAMED INDIVIDUAL` == "CHRISTOPHER CHADWICK/LC3/PORT CLYDE" ~ "C.C.",
      `NAMED INDIVIDUAL` == "DAVID TALOR/LC3/CUSHING" ~ "D.T.",
      `NAMED INDIVIDUAL` == "ERICH CULVER/LC2/PORT CLYDE" ~ "E.C.",
      `NAMED INDIVIDUAL` == "WHITNEY HUPPER/LC1/PORT CLYDE" ~ "W.H.",
      `NAMED INDIVIDUAL` == "GREG MORRIS/LC3/PORT CLYDE" ~ "G.M.",
      `NAMED INDIVIDUAL` == "JUSTIN THOMPSON/LC3/PORT CLYDE" ~ "J.T.",
      `NAMED INDIVIDUAL` == "DANIEL MORRIS/LC2/PORT CLYDE" ~ "D.M.",
      `NAMED INDIVIDUAL` == "DILLAN CUSHMAN/LC3/PORT CLYDE" ~ "D.C.",
      `NAMED INDIVIDUAL` == "SHANE HATCH/LC3/OWLS HEAD" ~ "S.H.",
      `NAMED INDIVIDUAL` == "WINSTON PEASE/LCO/PORT CLYDE" ~ "W.P.",
      `NAMED INDIVIDUAL` == "CHARLES CHRISTENSEN/LC1/PORT CLYDE" ~ "C.C.",
      `NAMED INDIVIDUAL` == "ANTHONY HOOPER/LC3/TENANTS HBR" ~ "A.H.",
      `NAMED INDIVIDUAL` == "BRIAN YORK/LC2/PORT CLYDE" ~ "B.Y.",
      `NAMED INDIVIDUAL` == "DYLAN LORD/STUDENT/ST GEORGE" ~ "D.L.",
      `NAMED INDIVIDUAL` == "JADEN PETERSDORF/LC2/PORT CLYDE" ~ "J.P.",
      `NAMED INDIVIDUAL` == "6691 IS LCS IN 2021; THESE TAGS ARE TOO OLD TO BE LCS" ~ "N/A",
      `NAMED INDIVIDUAL` == "NOT IN 2021 DMR LICENSE LIST" ~ "N/A"
    )) |>
  select(-`NAMED INDIVIDUAL`, -c(17:27), -`MISCELLANEOUS`)

write.csv(lobsters_anon, "../data/lobsters_anon.csv", row.names = FALSE)
```

## 4. Data Ethics Review

The data ethics review section will be introduced in a separate class and is not part of the original proposal deadline.



```{r homeport_points}
library(leaflet)

# Define your locations
locations <- data.frame(
  name = c("PORT CLYDE", "WINTER HBR", "ROCKLAND", "TENANTS HBR", "CUSHING", "OWLS HEAD", "ST GEORGE"),
  lat = c(43.9271, 44.3948, 44.1037, 43.9682, 44.0138, 44.0823, 44.0165),
  lon = c(-69.2583, -68.0830, -69.1089, -69.2088, -69.2615, -69.0573, -69.1989)
)

target_lat <- 43.90538
target_lon <- -69.266659


leaflet(data = locations) %>%
  addTiles() %>%
  addMarkers(~lon, ~lat, popup = ~name) %>%
  addMarkers(lng = -69.266659, lat = 43.90538, popup = "HART ISLAND")
```


```{r homeport_maps}
```


```{r homeport_maps}
# Step 1: Summarize trap counts by homeport
homeport_counts <- lobsters_anon |>
  group_by(`HOME PORT`) |>
  summarise(
    trap_count = n(),
    Longitude = first(`HOMEPORT LONG`),
    Latitude = first(`HOMEPORT LAT`)
  ) 
locations <- data.frame(
  name = c("PORT CLYDE", "WINTER HBR", "ROCKLAND", "TENANTS HBR", "CUSHING", "OWLS HEAD", "ST GEORGE"),
  lat = c(43.9271, 44.3948, 44.1037, 43.9682, 44.0138, 44.0823, 44.0165),
  lon = c(-69.2583, -68.0830, -69.1089, -69.2088, -69.2615, -69.0573, -69.1989)
)

target_lat <- 43.90538
target_lon <- -69.266659

# Filter out rows where homeport coordinates are NA
homeports_known <- homeport_counts |>
  filter(!is.na(Longitude) & !is.na(Latitude))

# Step 2: Set coordinates for Hart Island, ensuring they are numeric
hart_island <- c(
  longitude = as.numeric(unique(lobsters_anon$`GPS COORD-Long`)[1]),
  latitude = as.numeric(unique(lobsters_anon$`GPS COORDINATES-Lat`)[1])
)

# Step 3: Create spatial lines for each homeport leading to Hart Island
# Convert to an sf object with lines for each homeport to Hart Island
lines <- homeports_known |>
  rowwise() |>
  mutate(
    geometry = st_sfc(
      st_linestring(
        matrix(
          c(Longitude, Latitude, hart_island["longitude"], hart_island["latitude"]),
          ncol = 2, byrow = TRUE
        )
      )
    )
  ) |>
  st_as_sf(crs = 4326)

# Step 4: Define a color scale for markers
color_scale <- colorNumeric(palette = "YlOrRd", domain = homeport_counts$trap_count)

# Define your locations
locations <- data.frame(
  name = c("PORT CLYDE", "WINTER HBR", "ROCKLAND", "TENANTS HBR", "CUSHING", "OWLS HEAD", "ST GEORGE"),
  lat = c(43.9271, 44.3948, 44.1037, 43.9682, 44.0138, 44.0823, 44.0165),
  lon = c(-69.2583, -68.0830, -69.1089, -69.2088, -69.2615, -69.0573, -69.1989)
)

target_lat <- 43.90538
target_lon <- -69.266659

# Create a color scale for trap counts
color_scale <- colorNumeric(
  palette = "YlOrRd", # Color palette (Yellow-Orange-Red)
  domain = homeport_counts$trap_count, # Range of values
  na.color = "gray" # Handle missing values
)

# Step 5: Create the leaflet map
leaflet() |>
  addProviderTiles("Stadia.AlidadeSmoothDark") |>
  # Add lines with thickness proportional to trap count
  addPolylines(
    data = lines,
    weight = 3,
    color = "blue",
    opacity = 0.7,
    label = ~paste(`HOME PORT`, "to Hart Island:", trap_count, "traps")
  ) |>
  # Add a marker for Hart Island
  addCircleMarkers(
    lng = hart_island["longitude"],
    lat = hart_island["latitude"],
    color = "green",
    radius = 5,
    label = "Hart Island"
  ) |>
  # Add markers for homeports with colors based on trap_count
  addCircleMarkers(
    data = homeport_counts,
    lng = ~Longitude,
    lat = ~Latitude,
    radius = 8,
    color = ~color_scale(trap_count),  # Set border color
    fillColor = ~color_scale(trap_count),  # Set fill color
    fillOpacity = 0.8,  # Adjust opacity for visibility
    label = ~paste(`HOME PORT`, ": ", trap_count, "traps")
  ) |>
  
  # Add legend
  addLegend(
    position = "bottomright",
    pal = color_scale,
    values = homeport_counts$trap_count,
    title = "Trap Count",
    opacity = 0.7
  )
```
<<<<<<< HEAD



=======
>>>>>>> 34f4281e4229358b4340b3ec939910c37f70c098
