---
title: "Hart Island Lobster Traps"
author: "Big Harts"
output: html_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)

```

## 1. Introduction

This data set was provided from the Rozalia Project, a non-profit that works on collection, research, and education of marine debis around the Gulf of Maine. Most of what debris is collected and cataloged is tied to the fishing industries in the region.

This data was collected on August 24th, 2023 on Hart Island, a small island just
Southwest of Port Clyde, ME. Volunteers worked to  removing lost, abandoned and derelict fishing gear and other coastal debris. For each lobster trap found, the trap tags were pulled and recorded including: tag count, data collected (MDY), number of tags in bundle, state, zone, area, tag ID, trap batch number, tag color, shoreline where found, GPS (lat/long), tag year, additional info on tag, named individual, and home port. 

With this data we aim to explore the relationship between abandoned lobster traps and the individuals who deployed them and the lifespan of traps. 

## 2. Data

```{r}
lobsters <- read_csv(file = "../data/lobsters.csv")
glimpse(lobsters)
```

## 3. Data analysis plan

Text goes here.

3) the prelim_vis_testing graph below helps show the number of lost traps per individual, and the age_traps graph shows the correlation between age of traps and number of traps

```{r}
4) we hope to build on the graphs below to better show relationships between traps and individuals and show the range of trap age. 

```{r prelim_vis_testing}
ggplot(data = lobsters, mapping = aes(x = `NAMED INDIVIDUAL`)) +
  geom_bar() 


```
```{r age_traps}
ggplot(data = lobsters, mapping = aes(x = `# TAGS IN BUNDLE`)) +
  geom_bar()
```


## 4. Data Ethics Review

The data ethics review section will be introduced in a separate class and is not part of the original proposal deadline.

