---
title: "Hart Isalnd Trap memo"
author: "Big Harts!"
output: github_document
---

This document should contain a detailed account of the data clean up for your data and the design choices you are making for your plots. For instance you will want to document choices you've made that were intentional for your graphic, e.g. color you've chosen for the plot. Think of this document as a code script someone can follow to reproduce the data cleaning steps and graphics in your handout.

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
```

## Data Clean Up Steps for Overall Data

### Step 1: Load original dataset

```{r lobsters}
lobsters <- read_csv(file = "../data/lobsters.csv")
glimpse(lobsters)
```

### Step 2: Data cleaning

Remove unnecessary text rows at bottom of dataset:
```{r true_lobsters}
true_lobsters <- head(lobsters, -4)
glimpse(true_lobsters)
```

Add in coordinates to each homeport:
```{r homeport_coords}
homeport_coords <- true_lobsters |>
  mutate(
    "HOMEPORT LAT" = case_when(
     `HOME PORT` == "PORT CLYDE" ~ 43.9271,
     `HOME PORT` == "WINTER HBR" ~ 44.3948,
     `HOME PORT` == "ROCKLAND" ~ 44.1037,
     `HOME PORT` == "TENANTS HBR" ~ 43.9682,
     `HOME PORT` == "CUSHING" ~ 44.0138,
     `HOME PORT` == "OWLS HEAD" ~ 44.0823,
     `HOME PORT` == "ST GEORGE" ~ 44.0165
    )) |>
  mutate(
    "HOMEPORT LONG" = case_when(
     `HOME PORT` == "PORT CLYDE" ~ -69.2583,
     `HOME PORT` == "WINTER HBR" ~ -68.0830,
     `HOME PORT` == "ROCKLAND" ~ -69.1089,
     `HOME PORT` == "TENANTS HBR" ~ -69.2088,
     `HOME PORT` == "CUSHING" ~ -69.2615,
     `HOME PORT` == "OWLS HEAD" ~ -69.0573,
     `HOME PORT` == "ST GEORGE" ~ -69.1989
    ))
```

Assign acronyms to each individual and then remove their names to anonymise them:
```{r anon_data}
lobsters_anon <- homeport_coords |>   
  mutate(
    "INDIVIDUALS" = case_when(
      `NAMED INDIVIDUAL` == "THOMAS YOUNG/LC2O/WINTER HBR" ~ "T.Y.",
      `NAMED INDIVIDUAL` == "DOUGLAS ANDERSON/LC3/PORT CLYDE" ~ "D.A.",
      `NAMED INDIVIDUAL` == "KEITH YORK/LC2/ROCKLAND" ~ "K.Y.",
      `NAMED INDIVIDUAL` == "PETER HENDERSON/LNC/TENANTS HBR" ~ "P.H.",                      
      `NAMED INDIVIDUAL` == "CHRISTOPHER ANDERSON/LC2/PORT CLYDE" ~ "C.A.",                  
      `NAMED INDIVIDUAL` == "GARY LIBBY/LC2/PORT CLYDE" ~ "G.L.",                            
      `NAMED INDIVIDUAL` == "CHRISTOPHER CHADWICK/LC3/PORT CLYDE" ~ "C.C.",
      `NAMED INDIVIDUAL` == "DAVID TALOR/LC3/CUSHING" ~ "D.T.",
      `NAMED INDIVIDUAL` == "ERICH CULVER/LC2/PORT CLYDE" ~ "E.C.",
      `NAMED INDIVIDUAL` == "WHITNEY HUPPER/LC1/PORT CLYDE" ~ "W.H.",
      `NAMED INDIVIDUAL` == "GREG MORRIS/LC3/PORT CLYDE" ~ "G.M.",
      `NAMED INDIVIDUAL` == "JUSTIN THOMPSON/LC3/PORT CLYDE" ~ "J.T.",
      `NAMED INDIVIDUAL` == "DANIEL MORRIS/LC2/PORT CLYDE" ~ "D.M.",
      `NAMED INDIVIDUAL` == "DILLAN CUSHMAN/LC3/PORT CLYDE" ~ "D.C.",
      `NAMED INDIVIDUAL` == "SHANE HATCH/LC3/OWLS HEAD" ~ "S.H.",
      `NAMED INDIVIDUAL` == "WINSTON PEASE/LCO/PORT CLYDE" ~ "W.P.",
      `NAMED INDIVIDUAL` == "CHARLES CHRISTENSEN/LC1/PORT CLYDE" ~ "C.C.",
      `NAMED INDIVIDUAL` == "ANTHONY HOOPER/LC3/TENANTS HBR" ~ "A.H.",
      `NAMED INDIVIDUAL` == "BRIAN YORK/LC2/PORT CLYDE" ~ "B.Y.",
      `NAMED INDIVIDUAL` == "DYLAN LORD/STUDENT/ST GEORGE" ~ "D.L.",
      `NAMED INDIVIDUAL` == "JADEN PETERSDORF/LC2/PORT CLYDE" ~ "J.P.",
      `NAMED INDIVIDUAL` == "6691 IS LCS IN 2021; THESE TAGS ARE TOO OLD TO BE LCS" ~ "N/A",
      `NAMED INDIVIDUAL` == "NOT IN 2021 DMR LICENSE LIST" ~ "N/A"
    )) |>
  select(-`NAMED INDIVIDUAL`, -c(17:27), -`MISCELLANEOUS`)
```
## Plots


### Plot 1: Preliminary data visualisation

```{r prelim_vis_testing, fig.alt= "Bar plot of lobster traps collected on Hart Island that shows the number of traps belonging to each individual. The plot tells us that Brian York had the most traps, over 80"}
ggplot(data = true_lobsters, mapping = aes(x = `NAMED INDIVIDUAL`)) +
  geom_bar() +
theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

### Plot 2: Test plot of trap ages

```{r age_graph, fig.alt= "Bar plot showing the ages of lobster traps collected on Hart Island, displaying the amount of traps at each age. 1 year of age has the highest amount, with trap count decreasing as age increases, showing that fewer traps survive as time passes"}
ggplot(data = true_lobsters, mapping = aes(x = `# TAGS IN BUNDLE`)) +
  geom_bar() +
  labs(x = "Age of Traps (Number of Tags in Bundle)", y = "Number of Traps",
  title = "Age of Traps Collected on Hart Island") +
  theme_linedraw() +
  scale_x_continuous(breaks = seq(0, 10, by = 1))
```


### Plot 3: Trap IDs Found on Hart Island

<<<<<<< HEAD
#### Data cleanup steps specific to plot 3
This graph is showing the correlation of number of lobster traps to individuals and their home ports. We can see the majority of traps found came from Port Clyde. While most of the individuals are linked to 30 or less traps, one outlier from Port Clyde has closer to 85 lost traps. To tidy the data, we arranged the graph in descending order using arrange(desc(n)) function, and mutated the N/A variable to read as "Unknown" using the mutate() function. 
=======
#### Data Cleanup
>>>>>>> 30c83b247dbfafe22eaf92c1cbb4f81c211dc8bd

```{r data_cleanup_p3}
  trap_data <- true_lobsters %>%
  count(`TAG ID NUMBER`, `HOME PORT`) %>%
  arrange(desc(n)) %>%
  mutate(`HOME PORT` = if_else(`HOME PORT` == "N/A", "UNKNOWN", `HOME PORT`))

trap_data$`TAG ID NUMBER` <- fct_reorder(trap_data$`TAG ID NUMBER`, trap_data$n, .desc = TRUE)
```

#### Final Plot

```{r trap_ids, fig.alt= "Bar plot showing the number of lobster traps belonging to each individual collected on Hart Island. The plot shows that an overwhelming majority belonged to one person, with the ID 9899, having over 80 lost traps"}
ggplot(data = trap_data, aes(x = `TAG ID NUMBER`, y = n, fill = `HOME PORT`)) +
  geom_bar(stat = "identity") +
  labs(x = "Tag ID Number", y = "Number of Traps",
       title = "Trap IDs Found on Hart Island") +
  theme_linedraw() +
  theme(axis.text.x = element_text(angle = 70, hjust = 1)) +
  scale_fill_viridis_d(option = "H", name = "Home Port", direction = -1)
```

### Plot 4: Identify locations of each homeport on a map


#### Data Cleanup

```{r locations_dataset}
locations <- data.frame(
  name = c("PORT CLYDE", "WINTER HBR", "ROCKLAND", "TENANTS HBR", "CUSHING", "OWLS HEAD", "ST GEORGE"),
  lat = c(43.9271, 44.3948, 44.1037, 43.9682, 44.0138, 44.0823, 44.0165),
  lon = c(-69.2583, -68.0830, -69.1089, -69.2088, -69.2615, -69.0573, -69.1989)
)

target_lat <- 43.90538
target_lon <- -69.266659
```

#### Final Plot

```{r homeport_points, fig.alt= "Map with blue markers pointing to each of the home ports identified as an origin for at least one of the traps collected on Hart Island"}
library(leaflet)
leaflet(data = locations) %>%
  addTiles() %>%
  addMarkers(~lon, ~lat, popup = ~name) %>%
  addMarkers(lng = -69.266659, lat = 43.90538, popup = "HART ISLAND")
```

### Plot 5: 

#### Data Cleanup

```{r Map_setup}
#Summarise trap counts by homeport
homeport_counts <- lobsters_anon |>
  group_by(`HOME PORT`) |>
  summarise(
    trap_count = n(),
    Longitude = first(`HOMEPORT LONG`),
    Latitude = first(`HOMEPORT LAT`)
  ) 

#Filter out NA values
homeports_known <- homeport_counts |>
  filter(!is.na(Longitude) & !is.na(Latitude))

#Set Hart Island coordinates
hart_island <- c(
  longitude = as.numeric(unique(lobsters_anon$`GPS COORD-Long`)[1]),
  latitude = as.numeric(unique(lobsters_anon$`GPS COORDINATES-Lat`)[1])
)

#Create lines linking each home port to Hart Island
lines <- homeports_known |>
  rowwise() |>
  mutate(
    geometry = st_sfc(
      st_linestring(
        matrix(
          c(Longitude, Latitude, hart_island["longitude"], hart_island["latitude"]),
          ncol = 2, byrow = TRUE
        )
      )
    )
  ) |>
  st_as_sf(crs = 4326)

#Home port colour scale
color_scale <- colorNumeric(palette = "YlOrRd", domain = homeport_counts$trap_count)
```

#### Final Plot
```{r homeport_maps, fig.alt= "Map with markers showing each of the identified home ports that at least one trap was identified as originating from. Home port markers are coloured by the number of traps that came from them, with blue lines linking each home port to Hart Island, where the traps were collected"}
leaflet() |>
  addProviderTiles("Stadia.AlidadeSmoothDark") |>
  
  #Homeport Lines
  addPolylines(
    data = lines,
    weight = 3,
    color = "blue",
    opacity = 0.7,
    label = ~paste(`HOME PORT`, "to Hart Island:", trap_count, "traps")
  ) |>
  
  #Hart Island Marker
  addCircleMarkers(
    lng = hart_island["longitude"],
    lat = hart_island["latitude"],
    color = "green",
    radius = 5,
    label = "Hart Island"
  ) |>
  
  #Homeport Markers
  addCircleMarkers(
    data = homeport_counts,
    lng = ~Longitude,
    lat = ~Latitude,
    radius = 8,
    color = ~color_scale(trap_count),  # Set border color
    fillColor = ~color_scale(trap_count),  # Set fill color
    fillOpacity = 0.8,  # Adjust opacity for visibility
    label = ~paste(`HOME PORT`, ": ", trap_count, "traps")
  ) |>
  
  # Map legend
  addLegend(
    position = "bottomright",
    pal = color_scale,
    values = homeport_counts$trap_count,
    title = "Trap Count",
    opacity = 0.7
  )
```