---
title: "Hart Isalnd Trap memo"
author: "Big Harts"
output: github_document
---

This document should contain a detailed account of the data clean up for your data and the design choices you are making for your plots. For instance you will want to document choices you've made that were intentional for your graphic, e.g. color you've chosen for the plot. Think of this document as a code script someone can follow to reproduce the data cleaning steps and graphics in your handout.

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
```

## Data Clean Up Steps for Overall Data

### Step 1: Load original dataset

```{r lobsters}
lobsters <- read_csv(file = "../data/lobsters.csv")
glimpse(lobsters)
```

### Step 2: Data cleaning

Remove unnecessary text rows at bottom of dataset:
```{r true_lobsters}
true_lobsters <- head(lobsters, -4)
glimpse(true_lobsters)
```

Add in coordinates to each homeport:
```{r homeport_coords}
homeport_coords <- true_lobsters |>
  mutate(
    "LATITUDE" = case_when(
     `HOME PORT` == "PORT CLYDE" ~ 43.9271,
     `HOME PORT` == "WINTER HBR" ~ 44.3948,
     `HOME PORT` == "ROCKLAND" ~ 44.1037,
     `HOME PORT` == "TENANTS HBR" ~ 43.9682,
     `HOME PORT` == "CUSHING" ~ 44.0138,
     `HOME PORT` == "OWLS HEAD" ~ 44.0823,
     `HOME PORT` == "ST GEORGE" ~ 44.0165
    )) |>
  mutate(
    "LONGITUDE" = case_when(
     `HOME PORT` == "PORT CLYDE" ~ -69.2583,
     `HOME PORT` == "WINTER HBR" ~ -68.0830,
     `HOME PORT` == "ROCKLAND" ~ -69.1089,
     `HOME PORT` == "TENANTS HBR" ~ -69.2088,
     `HOME PORT` == "CUSHING" ~ -69.2615,
     `HOME PORT` == "OWLS HEAD" ~ -69.0573,
     `HOME PORT` == "ST GEORGE" ~ -69.1989
    ))
```

Assign acronyms to each individual and then remove their names to anonymise them:
```{r anon_data}
lobsters_anon <- homeport_coords |>   
  mutate(
    "INDIVIDUALS" = case_when(
      `NAMED INDIVIDUAL` == "THOMAS YOUNG/LC2O/WINTER HBR" ~ "T.Y.",
      `NAMED INDIVIDUAL` == "DOUGLAS ANDERSON/LC3/PORT CLYDE" ~ "D.A.",
      `NAMED INDIVIDUAL` == "KEITH YORK/LC2/ROCKLAND" ~ "K.Y.",
      `NAMED INDIVIDUAL` == "PETER HENDERSON/LNC/TENANTS HBR" ~ "P.H.",                      
      `NAMED INDIVIDUAL` == "CHRISTOPHER ANDERSON/LC2/PORT CLYDE" ~ "C.A.",                  
      `NAMED INDIVIDUAL` == "GARY LIBBY/LC2/PORT CLYDE" ~ "G.L.",                            
      `NAMED INDIVIDUAL` == "CHRISTOPHER CHADWICK/LC3/PORT CLYDE" ~ "C.C.",
      `NAMED INDIVIDUAL` == "DAVID TALOR/LC3/CUSHING" ~ "D.T.",
      `NAMED INDIVIDUAL` == "ERICH CULVER/LC2/PORT CLYDE" ~ "E.C.",
      `NAMED INDIVIDUAL` == "WHITNEY HUPPER/LC1/PORT CLYDE" ~ "W.H.",
      `NAMED INDIVIDUAL` == "GREG MORRIS/LC3/PORT CLYDE" ~ "G.M.",
      `NAMED INDIVIDUAL` == "JUSTIN THOMPSON/LC3/PORT CLYDE" ~ "J.T.",
      `NAMED INDIVIDUAL` == "DANIEL MORRIS/LC2/PORT CLYDE" ~ "D.M.",
      `NAMED INDIVIDUAL` == "DILLAN CUSHMAN/LC3/PORT CLYDE" ~ "D.C.",
      `NAMED INDIVIDUAL` == "SHANE HATCH/LC3/OWLS HEAD" ~ "S.H.",
      `NAMED INDIVIDUAL` == "WINSTON PEASE/LCO/PORT CLYDE" ~ "W.P.",
      `NAMED INDIVIDUAL` == "CHARLES CHRISTENSEN/LC1/PORT CLYDE" ~ "C.C.",
      `NAMED INDIVIDUAL` == "ANTHONY HOOPER/LC3/TENANTS HBR" ~ "A.H.",
      `NAMED INDIVIDUAL` == "BRIAN YORK/LC2/PORT CLYDE" ~ "B.Y.",
      `NAMED INDIVIDUAL` == "DYLAN LORD/STUDENT/ST GEORGE" ~ "D.L.",
      `NAMED INDIVIDUAL` == "JADEN PETERSDORF/LC2/PORT CLYDE" ~ "J.P.",
      `NAMED INDIVIDUAL` == "6691 IS LCS IN 2021; THESE TAGS ARE TOO OLD TO BE LCS" ~ "N/A",
      `NAMED INDIVIDUAL` == "NOT IN 2021 DMR LICENSE LIST" ~ "N/A"
    )) |>
  select(-`NAMED INDIVIDUAL`, -c(17:27), -`MISCELLANEOUS`)
```
## Plots

### ggsave example for saving plots

```{r starwars ggsave example}
p1 <- starwars |>
  filter(mass < 1000, 
         species %in% c("Human", "Cerean", "Pau'an", "Droid", "Gungan")) |>
  ggplot() +
  geom_point(aes(x = mass, 
                 y = height, 
                 color = species)) +
  labs(x = "Weight (kg)", 
       y = "Height (m)",
       color = "Species",
       title = "Weight and Height of Select Starwars Species",
       caption = paste("This data comes from the starwars api: https://swapi.py43.com"))


ggsave("example-starwars.png", width = 4, height = 4)

ggsave("example-starwars-wide.png", width = 6, height = 4)
```



### Plot 1: _________

```{r traps_ages_1}
ggplot(data = true_lobsters, mapping = aes(x = `# TAGS IN BUNDLE`)) +
  geom_bar() +
  labs(x = "Age of Traps (Years)", y = "Number of Traps",
  title = "Age of Traps Collected on Hart Island") +
  theme_linedraw() +
  scale_x_reverse(breaks = seq(10, 0, by = -1))

ggsave("hart-traps-ages.png", width = 4, height = 4)
```

```{r 2_age_graph}
ggplot(data = true_lobsters, mapping = aes(x = `# TAGS IN BUNDLE`)) +
  geom_bar() +
  labs(x = "Age of Traps (Years)", y = "Number of Traps",
  title = "Age of Traps Collected on Hart Island") +
  theme_linedraw() +
  scale_x_continuous(breaks = seq(0, 10, by = 1))
  
ggsave("hart-traps-ages-2.png", width = 4, height = 4)
```

#### Data cleanup steps specific to plot 1

These data cleaning sections are optional and depend on if you have some data cleaning steps specific to a particular plot

```{r}

```


#### Final Plot 1

```{r}

```


### Plot 2: _________

### Plot 3: ___________

Add more plot sections as needed. Each project should have at least 3 plots, but talk to me if you have fewer than 3.

### Plot 4: ___________ 

